Explore and Summarize white win quality by abdulazez alsulaiman 
========================================================
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(knitr)
library(dplyr)
library(corrplot)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data <- read.csv("wineQualityWhites.csv", header=TRUE, sep=",")

```
this dataset of white win quality, it has 4898 columns,
what is intesesting about it is varias compenation in a qlass of win,
in this report i am going to explore this data and decover wondufel things.


```{r echo=FALSE, discover_the_Data}
# discover the data 
str(data)
summary(data)
```
```{r echo=FALSE, clean_the_data}
# Delet the first column because its usles and make noise in the data 
data$X <- NULL
str(data)
```
# Univariate Plots Section

```{r echo=FALSE, build_one_function}
# build function for one variable plot to avoid defulction in code 
plot_one_variable <- function(xa , m , xl){
plot <-ggplot(data,aes(x = xa ))+ geom_histogram ()+labs(title=m, x=xl)
return(plot) 
}

```

```{r echo=FALSE, Univariate_Plots1}
#call the one variable plot for all columns in the data 
plot_one_variable(data$fixed.acidity , "histogram for fixed.acidity" ,
                  "fixed acidity (tartaric acid - g / dm^3)")

```
as we can see in the apove graph the data is normaly distributed between 5 to 9
 

```{r echo=FALSE, Univariate_Plots2}
#call the one variable plot for all columns in the data 

plot_one_variable(data$volatile.acidity , "histogram for volatile.acidity" ,
                  "volatile acidity (acetic acid - g / dm^3)")

```

 
as we can see in the apove graph the most tested data have low volatile acidity 
 
```{r echo=FALSE, Univariate_Plots3}
#call the one variable plot for all columns in the data 

plot_one_variable(data$citric.acid, "histogram for citric.acid" ,
                  "citric acid (g / dm^3)")

```


as we can see in the apove graph the data is normaly distributed between 0 to 0.5

```{r echo=FALSE, Univariate_Plots4}
#call the one variable plot for all columns in the data 

plot_one_variable(data$residual.sugar , "histogram for residual.sugar" , 
                  "residual sugar (g / dm^3)")


```

as we can see in the apove graph
most of tested data have residual sugar from 4 to 8

```{r echo=FALSE, Univariate_Plots5}
#call the one variable plot for all columns in the data 

plot_one_variable(data$free.sulfur.dioxide ,
                  "histogram for  free.sulfur.dioxide" ,
                  " free sulfur dioxide (mg / dm^3)")


```

as we can see in the apove graphthe free sulfur dioxide is normaly distributed

```{r echo=FALSE, Univariate_Plots6}
#call the one variable plot for all columns in the data 

plot_one_variable(data$chlorides , "histogram for chlorides" ,
                  "chlorides (sodium chloride - g / dm^3)")


```

as shown in the graph 6 the most data have chlorides between 0 to 0.1

```{r echo=FALSE, Univariate_Plots7}
#call the one variable plot for all columns in the data 

plot_one_variable(data$total.sulfur.dioxide ,
                  "histogram for total.sulfur.dioxide" ,
                  "total sulfur dioxide (mg / dm^3)")

```


as shown in the apove graph the total sulfur dioxide is normaly distributed


```{r echo=FALSE, Univariate_Plots8}
#call the one variable plot for all columns in the data 

plot_one_variable(data$density , "histogram for density" , "density (g / cm^3)")

```


as we can see in the apove graph the most data have density lower than 1

```{r echo=FALSE, Univariate_Plots9}
#call the one variable plot for all columns in the data 

plot_one_variable(data$pH , "histogram for pH" , "pH")

```

as we can see in the apove graph the pH data is normaly distributed

```{r echo=FALSE, Univariate_Plots10}
#call the one variable plot for all columns in the data 

plot_one_variable(data$sulphates , "histogram for sulphates" ,
                  "sulphates (potassium sulphate - g / dm3)")

```

as we can see in the apove graph 
the most data have sulphates between 0.3 and 0.6

```{r echo=FALSE, Univariate_Plots11}
#call the one variable plot for all columns in the data 

plot_one_variable(data$alcohol , "histogram for alcohol" ,
                  "alcohol (% by volume)")

```

as we can see in the apove graph 
there is few data have higher alcohol percentage 

```{r echo=FALSE, Univariate_Plots12}
#call the one variable plot for all columns in the data 

plot_one_variable(data$quality , "histogram for quality" ,
                  "quality (score between 0 and 10)")

```


as shown in the apove graphmost
most common quality in tha data is 6, over 2000 samples have it.







# Univariate Analysis


### What is the structure of your dataset?
there is 13 variables in my dataset and 4898 rwos 

### What is/are the main feature(s) of interest in your dataset?
1- quality , 2-pH , 3-alcohol , 4-density , 5-residual.sugar

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

i think thes features is ok for me 
### Did you create any new variables from existing variables in the dataset?
no i did not 
### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

i delet column X becuase it makes noise in the data and its useless 

# Bivariate Plots Section



```{r echo=FALSE, Bivariate_Plots1}
#plot matrix correlation to see the correlation
#between the variables in the data 
core <-cor(data)
corrplot(core, method='number'  , mar=c(0,0,1,0),
         title="matrix correlation for all the features in numbers")

```

the apove graph shows the matrix correlation for all the features in numbers


```{r echo=FALSE, Bivariate_Plots2}
#plot matrix correlation to see the correlation
#between the variables in the data 
corrplot(core, method='square',mar=c(0,0,1,0),
         title="matrix correlation for all the features in square") 

```

the apove graph shows the matrix correlation for all the features in squares


```{r echo=FALSE, build_two_function}
#build function for two variables plot to avoid defluction in code 
two_variables_plot<-function(x_axis , y_axis ,m ,xl , yl){
plot<-ggplot(data, aes(x=x_axis, y= y_axis )) + geom_point() + 
geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95) + 
labs(title=m ,x=xl ,y = yl)
return(plot)
}

```

```{r echo=FALSE, Bivariate_Plots3}
#call the two variables to see the relationships between diffirent variables 
two_variables_plot(data$density , data$residual.sugar,
                   "the correlation between density and residual.sugar " ,
                   "density (g / cm^3) ", "residual sugar (g / dm^3) ")


```

in the apove graph we can see the the relationship
between density and residual sugar 

```{r echo=FALSE, Bivariate_Plots4}
#call the two variables to see the relationships between diffirent variables 

two_variables_plot(data$alcohol , data$density ,
                   "the correlation between alcohol and density " ,
                   "alcohol (% by volume) ", "density (g / cm^3)")

```

in the apove graph we can see the relationship between alcohol and density 


```{r echo=FALSE, Bivariate_Plots5}
#call the two variables to see the relationships between diffirent variables 

two_variables_plot(data$free.sulfur.dioxide,data$residual.sugar,
                   "the correlation between free.sulfur.dioxide
                   and residual.sugar","free sulfur dioxide (mg / dm^3)",
                   "total sulfur dioxide (mg / dm^3) ")


```



in the apove graph we can see the relationship between free sulfur dioxide
and residual.sugar


```{r echo=FALSE, Bivariate_Plots6}
#call the two variables to see the relationships between diffirent variables 

two_variables_plot(data$fixed.acidity , data$pH ,
                   "the correlation between fixed.acidity and pH " ,
                   "fixed acidity (tartaric acid - g / dm^3) ", "pH")

```

in the apove graph we can see the relationship between fixed acidity and pH 



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

i see some positive correlation and negitive correlation .
### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
no because it will be useless 
### What was the strongest relationship you found?
relationship between density and residual.sugar 



# Multi variate Plots Section


```{r echo=FALSE, Multivariate_Plots1}
#plot the relationships between two variables and the therd is quality 

ggplot(aes(x = alcohol, y = density ), data = data) +
  geom_point(aes(color = factor(quality))) +
  scale_colour_brewer()+
  ggtitle("relationship between density and dalcohol and qulaity ")+
labs (x="alcohol (% by volume)" ,y ="density (g / cm^3)") 





```


in the apove graph we can see the relationship
between alcohol, density and quality 


```{r echo=FALSE, Multivariate_Plots2}
#plot the relationships between two variables and the therd is quality 



ggplot(aes(x=density, y= residual.sugar  ), data = data) +
  geom_point(aes(color = factor(quality))) +
  scale_colour_brewer()+
  ggtitle("relationship between density and residual.sugar and qulaity ")+
labs (x="residual sugar (g / dm^3)" ,y ="density (g / cm^3)")





```


in the apove graph we can see the the relationship
between density , residual sugar and quality 

```{r echo=FALSE, Multivariate_Plots3}
#plot the relationships between two variables and the therd is quality 


ggplot(aes(x= free.sulfur.dioxide, y= total.sulfur.dioxide  ), data = data) +
  geom_point(aes(color = factor(quality))) +
  scale_colour_brewer()+
  ggtitle("relationship between free.sulfur.dioxide 
          and total.sulfur.dioxide and qulaity ")+
labs (x="free sulfur dioxide (mg / dm^3)" , 
      y ="total sulfur dioxide (mg / dm^3)")



```

in the apove graph we can see relationship 
between free.sulfur.dioxide and total.sulfur.dioxide and qulaity 

```{r echo=FALSE, Multivariate_Plots4}
#plot the relationships between two variables and the therd is quality 


ggplot(aes(x=fixed.acidity, y=pH ), data = data) +
  geom_point(aes(color = factor(quality))) +
  scale_colour_brewer()+
  ggtitle("relationship between fixed.acidity and pH and qulaity ")+
labs (x="fixed acidity (tartaric acid - g / dm^3)" , 
      y ="pH")





```
in the apove graph we can see the relationship
between fixed acidity , pH and quality 



# multi varaible Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

i see the density has influnce in the data and has strong relationship with 
other features 
### Were there any interesting or surprising interactions between features?
ther qulaity increase when alcohol increase that is intersting 




# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}
plot_one_variable(data$pH , "histogram for pH" , "pH")

```

### Description One
as we can see in the apove graph the data in pH is noramily distrebuted 


### Plot Two
```{r echo=FALSE, Plot_Two}
two_variables_plot(data$fixed.acidity , data$pH ,
                   "the correlation between fixed.acidity and pH " ,
                   "fixed acidity (tartaric acid - g / dm^3) ", "pH")
```

### Description Two
as we can see thwe apove graph when the fixed.acidity is low the pH is higer 
### Plot Three
```{r echo=FALSE, Plot_Three}


ggplot(aes(x = alcohol, y = density ), data = data) +
  geom_point(aes(color = factor(quality))) +
  scale_colour_brewer()+
  ggtitle("relationship between density and dalcohol and qulaity ")+
labs (x="alcohol (% by volume)" ,y ="density (g / cm^3)") 


```



### Description Three
as we can see in the apove graph that if
the alcohol is higer and the densety is lower then the qulaity will be better 
------

# Reflection
struggles :
1- i don`t drink win and never see one , so it was not make sense 
2-the quality of the data is good but i had to make some chamges 

summary:
1-there is many positive and negative correlation between the data 
2- the most influnce factor in this data is density 
3-the most uninfunce factor is sulphates 


future work :

most of the tested win their qulaity is 6 and few 8 ,
it should be equal size of all diffirent qulaity level, 
so as a future idea, i think we need to collect more data with other qualities

